Gringo 2.0

### 2008-07-21 doc/documentation/paper.tex
dumb changes
git-svn-id: https://potassco.svn.sourceforge.net/svnroot/potassco/branches/gringo2@376 3e02f596-3131-0410-8352-e04eeaec37cc
example and explaination for writing logic programs
git-svn-id: https://potassco.svn.sourceforge.net/svnroot/potassco/branches/gringo2@406 3e02f596-3131-0410-8352-e04eeaec37cc

\subsection{Meta Constructs}
The ``\#'' sign in front of all meta expressions is optional and can therefore be omitted.
\paragraph{hide}
The hide construct is used to hide atoms from the output of the ASP solver.
Two different syntaxes are supported by gringo.
The first 
\begin{verbatim}
 #hide name/arity.
\end{verbatim}
and the second
\begin{verbatim}
 #hide name(X1, ... , Xn).
\end{verbatim}
Both variants hide the predicate ``name''.
The arity of the predicate can be given explicit or implicit with a list of variables.
The variables are just placeholder and do not unify. The second syntax is implemented for compatibility reasons only.
If the predicate that has to be hidden has arity zero, meaning it is an atom, the syntax is
\begin{verbatim}
 #hide name.
\end{verbatim}
Not to give explicitly all predicates that are hidden, only the ones which are visible the statement
\begin{verbatim}
 #hide.
\end{verbatim}
in conjunction with the ``show'' statement is used.
The ``show'' statement has the same syntax as ``hide'' but uses the keyword ``show''.
\begin{verbatim}
 #hide.
 #show p/3.
\end{verbatim}
The example hides all atoms except the predicate $p$ with arity $3$.
\paragraph{const}
Within the ``const'' statement, a default value for constants can be defined.
Constants are atoms used in the logic program and are normally defined via the command line options.
A default value for a constant $n$ can be easily be set.
\begin{verbatim}
 #const n=5.
\end{verbatim}
\paragraph{domain}
A domain statement states that a specific variable has a global domain restriction.
The statement
\begin{verbatim}
 #domain time(T).
\end{verbatim}
makes sure that the variable ``T'' is always bound to the domain of ``time''.
This has the same effect as adding the domain ``time(T)'' to every rule the variable occurs in.
\paragraph{compute}
The compute statement is a hint for the solver which models to compute.
Is has the syntax
\begin{verbatim}
 #compute number {a1, ..., aN, not aN+1, ..., not aM}.
\end{verbatim}
$number$ is the number of models that shall be computed by the answerset solver and can be omitted. If it is zero, no restriction to the number of models is applied.
All positive atoms in the compute statement must be in the computed models, and all negative atoms do not have to be in the models of the program.
The compute statement is solver dependent and supported by every solver.
\subsection{Features}
\paragraph{Expanding Terms}
\subparagraph{Range Operator}
To define several atoms that stretch on one or more parameters over a certain range, two dots can used to express an increasing integer range.
Using the range operator in the head of a rule, it works as a conjunction of all the members in the range, so for every number the rule is duplicated.
\begin{verbatim}
 times(1..3, 1..2).
\end{verbatim}
will result in the facts
\begin{verbatim}
 times(1,1). times(2,1). times(3,1).
 times(1,2). times(2,2). times(3,2).
\end{verbatim}

@@

\section{$\Pi$-Restrictiveness (or something else)}
\label{domainrestricted}
\section{Examples}
In this section come encodings for common problems are presented and explained.
\paragraph{Fastfood}
This example needs a proper instance to work.
\begin{verbatim}
#hide.
#show altdepot/2.
#show costs/1.

>>>>

\subsection{Meta Constructs}
The ``\#'' sign in front of all meta expressions is optional and can therefore be omitted.
\paragraph{hide}
The hide construct is used to hide atoms from the output of the ASP solver.
Two different syntaxes are supported by gringo.
The first 
\begin{verbatim}
 #hide name/arity.
\end{verbatim}
and the second
\begin{verbatim}
 #hide name(X1, ... , Xn).
\end{verbatim}
Both variants hide the predicate ``name''.
The arity of the predicate can be given explicit or implicit with a list of variables.
The variables are just placeholder and do not unify. The second syntax is implemented for compatibility reasons only.
If the predicate that has to be hidden has arity zero, meaning it is an atom, the syntax is
\begin{verbatim}
 #hide name.
\end{verbatim}
Not to give explicitly all predicates that are hidden, only the ones which are visible the statement
\begin{verbatim}
 #hide.
\end{verbatim}
in conjunction with the ``show'' statement is used.
The ``show'' statement has the same syntax as ``hide'' but uses the keyword ``show''.
\begin{verbatim}
 #hide.
 #show p/3.
\end{verbatim}
The example hides all atoms except the predicate $p$ with arity $3$.
\paragraph{const}
Within the ``const'' statement, a default value for constants can be defined.
Constants are atoms used in the logic program and are normally defined via the command line options.
A default value for a constant $n$ can be easily be set.
\begin{verbatim}
 #const n=5.
\end{verbatim}
\paragraph{domain}
A domain statement states that a specific variable has a global domain restriction.
The statement
\begin{verbatim}
 #domain time(T).
\end{verbatim}
makes sure that the variable ``T'' is always bound to the domain of ``time''.
This has the same effect as adding the domain ``time(T)'' to every rule the variable occurs in.
\paragraph{compute}
The compute statement is a hint for the solver which models to compute.
Is has the syntax
\begin{verbatim}
 #compute number {a1, ..., aN, not aN+1, ..., not aM}.
\end{verbatim}
$number$ is the number of models that shall be computed by the answerset solver and can be omitted. If it is zero, no restriction to the number of models is applied.
All positive atoms in the compute statement must be in the computed models, and all negative atoms do not have to be in the models of the program.
The compute statement is solver dependent and supported by every solver.

\subsection{How to write a logic program}
Writing a logic program to solve a problem is rather describing a problem than describing an algorithm to solve a problem.
Therefore the problem must be encoded as a set of rules.
Typically these are divided into a problem description and an instance of the problem.
The solution to a problem can then be read from answer sets of the rules.
The procedure of writing a logic program shall now be explained with a little example.
We want to write a problem encoding for the 3-colorability problem of graph coloring.
Given an arbitraty graph, no two connected edges may have the same color.
\newline
Usually you start with the non instance domains, like what colors we do have.
\begin{verbatim}
 color(red). color(green). color(blue).
\end{verbatim}
After that we have to color the vertices, but we must give attention that each vertex only has one color.
We can do this by simply saying that a vertex is colored with one color X, and that the same vertex does not have another color Y.
\begin{verbatim}
 colored(V,C) :- not colored(V,D), C != D, color(C;D), vertex(V).
\end{verbatim}
As you can see, you have to give the domain of the variables in the rule.
The variable V is used for vertices and the variables C and D are used for colors.
After every vertex now has a color we would get all possible solutions of all vertices colorfully painted.
We now need to restrict our solutions again by saying that two vertices that are connected by an edge should not have the same color.
We can do this by using an integrity constraint.
An integrity constraint is a rule with an empty head, and if an assignment of variables fulfills the body, the assignment is not a solution of our problem.
\begin{verbatim}
 :- edge(V,U), colored(V,C), colored(U,C), vertex(V;U), color(C).
\end{verbatim}
Here we have the edge between two vertices V and U, and if both are colored with the same color C the assignment is not a solution to our problem.
\newline
We have finally written our problem description, all we need now is a problem instance. This step is fairly simple.
We just need to define several vertices and edges that do connect the vertices.
\begin{verbatim}
 vertices(1..4).
 edge(1,2). edge(1,4). edge(2,4).
 edge(3,1). edge(4,1). edge(3,2).
\end{verbatim}

If we give the problem description and instance to the grounder, it will generate a variable free version of the problem. This can then be solved by a solver and in the answer sets of the solver the colored predicate states the color of a valid coloring of all vertices.
\newline
To ease the process writing logic programs some features in the input syntax were introduced and are now explained in detail.
\subsection{Features}
\paragraph{Expanding Terms}
\subparagraph{Range Operator}
To define several atoms that stretch on one or more parameters over a certain range, two dots can used to express an increasing integer range.
Using the range operator in the head of a rule, it works as a conjunction of all the members in the range, so for every number the rule is duplicated.
\begin{verbatim}
 times(1..3, 1..2).
\end{verbatim}
will result in the facts
\begin{verbatim}
 times(1,1). times(2,1). times(3,1).
 times(1,2). times(2,2). times(3,2).
\end{verbatim}

@@

\section{$\lambda$-Restriction}
\label{domainrestricted}
Gringo verifies that the input program is $\lambda$-
restricted \cite{gringopaper}, so that the existence of a finite equivalent ground instantiation is guaranteed. The definition of lambda restrictedness can be found in \cite{gringopaper} and was expanded for aggregates and disjunctions.

\section{Examples}
In this section come encodings for common problems are presented and explained.
\paragraph{Fastfood}
This example needs a proper instance to work.
\begin{verbatim}
#hide.
#show altdepot/2.
#show costs/1.
